---
title: "Using data.table the tidy way"
author: Huang Tian-Yuan (Hope)
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using data.table the tidy way}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Introduction


  For most functions, just use it like `tidyverse`, but add a `_dt` as the suffix. This is mostly true for verbs in `dplyr`. Nevertheless, the data.table has different features, and some API varies, but only varies for more convenience. Below are the supported verbs:
  
  
- basic manipulation: select_dt/slice_dt/filter_dt/mutate_dt/arrange_dt/summarise_dt/rename_dt/dist inct_dt
- count family: count_dt/add_count_dt/top_n_dt/top_frac_dt
- join: left_join_dt/right_join_dt/inner_join_dt/full_join_dt/anti_join_dt/semi_join_dt
- nest: nest_dt/unnese_dt
- wide-long transformation: longer_dt/wider_dt
- mising values: drop_na_dt/replace_na_dt
- fst: parse_fst/select_fst


I'll give only one example so far, just explore yourself and enjoy the fun!

```{r setup}
library(tidydt)

iris %>%
  mutate_dt(group = Species,sl = Sepal.Length,sw = Sepal.Width) %>%
  select_dt(group,sl,sw) %>%
  filter_dt(sl > 5) %>%
  arrange_dt(group,sl) %>%
  distinct_dt(sl,keep_all = T) %>%
  summarise_dt(sw = max(sw),by = group)
```

## Story 

   I am a big fan of `tidyverse`, however, when I have to deal with a data.frame of larger than 1G, I turned to `data.table`. But `data.table` is not easy, and is not so readable as `tidyverse`. Even today, I still need to explore for long time before I could use `data.table` correctly for specific occations. Therefore, when the data is smaller than 1G, I would always use `tidyverse`. If the file is between 1G and 10G, I would use `data.table` to do some major work and turned back to `tidyverse` when possible. This has been annoying acutually.  
   
  Until one day, I met `maditr` package. It told me that the manipulation of `data.table` does not have to be in this way. This could be done in a tidy way! And, you don't even need to use other packages to build this elegant API, you just need `data.table` and `magrittr`! I was shocked! To use a fast `tidyverse` is what I dream for long, and it could!  
  
   By now, Hadley has completed the `dtplyr`, which facilitate the workflow. However, many verbs are still not supported. I think it's my time to contribute. I used a lot of functions from `maditr`, but with a different API. It has a specific goal: provide users with state-of-the-art data manipulation tools with least pain and fastest speed.
